---
title: "Brownlows Bets"
author: "Rabi Elkhoury"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
source(file="BrLw.R")
options(digits=2)
```


<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 637 times so far. Where do you think his profits are? Discuss this in a markdown document
 
Report git commit structure:
There are 3 questions and there should be at least 3 separate commits, at least one dealing with each of the 3 questions.
1. What are the fair odds for a bet on matching birthdays in a 35 person class.
2. For what size class is a 2:1 odds bet fair.
3. How much money would Brownlow be expected to end up with if the bets went as described.

## Our strategy is to build a Function and the use it to answer the three questions. 

The function will be given a class_size by number of students, odds, and Number_of_times brownlow made the bet.  It will give you back how much Brownlows earnings were. We will use it to answer our three questions below. If you can, improve this function by giving default values to some of its inputs. 

# In the class and payoff as described  If you highered 40 brownlow substitute teachers to do the same thing that Brownlow did then.
## what is their average earnings

```{r}
Substitute <- numeric(length = 40)

for (j in 1:40) {
Substitute[j] = BrLw(odds = 2)
}
```


```{r}
mean(Substitute)
```
The average earnings, if 40 Brownlows were hired to play the game the exact same way, would be about `r mean(Substitute)` .


## the standard deviation and a 95% confidence interval for their earnings?

```{r}
sd(Substitute)
```
The standard deviation if 40 Brownlows were hired would be around `r sd(Substitute)`; if we consider the Central Limit Theorem, this would mean that a 95% confidence interval for the data would be within 2 standard deviations of the data, on both sides.  Thus, the 95% confidence interval ranges from `r mean(Substitute) - 2*sd(Substitute)` to `r mean(Substitute) + 2*sd(Substitute)` .
  
<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows winnings (or losings)-->


# What are the fair odds? Use manipulation of odds in the function to tune Brownlows earnings to zero. 

To determine fair odds, we are going to hire 1000 Brownlows to repeat the actions of the "real" Brownlow.
```{r}
Fair <- numeric(length = 1000)

for (j in 1:1000) {
Fair[j] = BrLw(odds = 4.38)
}
```

```{r}
hist(Fair)
```


We were able to find a reasonable interval centered at the fairest odds ("fair odds" imply the profit is zero), by testing with and generating an excess of Brownlow models.  The "fair odd" value was found to be 4.38, meaning that, if Brownlow were to wager $4.38 for every $1 given, there would be a profit of essentially zero.

<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 637 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->


# What size of class makes the 2:1 bet fair?  Use manipulation of number of Students to tune Brownlows earnings to zero.

<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting profits!-->

```{r}
Sub <- numeric(length = 1000)

for (j in 1:1000) {
Sub[j] = BrLw(numClass = 28)
}
```

```{r}
hist(Sub)
```

In running our simulations, it's become clear that a class size between 28 and 29 students is the number that'll make our 2:1 odd situation fair.  We can confirm this through plotting our simulation onto a histogram; for a class size of 28 students, most of the data is centered close to -20 in a bell-curve (indicating a Normal distribution).  Trying 29 for numClass yields a histogram centered around a positive number.  In actuality, fair odds probably require a numClass value that is a precise decimal.  Since we are dealing with numbers of students in a class (a discrete variable), we must round down to the smallest integer that preserves odds that are nearest to fair; as such, 28 students is the optimal answer.




# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->






